---
title: "ThetaProbe"
author: "Adam H. Sparks"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{ThetaProbe}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

An R package for internal use to retrieve files from the University of Southern
Queensland's NCEA soil probe FTP file server for the Centre for Crop Health's
Summer Crops Pathology group and collate all files while converting five-minute
data into daily values and filtering outliers in the five-minute data before
generating daily mean values for analysis and graphing.

## Load libraries for use in R session

```{r load_libraries, eval=FALSE}
library(ThetaProbe)
library(ggplot2)
library(scales)
library(readr)
library(theme.usq)
```

## Using get_os() to set the filepaths

Check the operating system first to see what platform R is running on. Set the
output file path appropriately.

```{r check_os, eval=FALSE}

if(get_os == "windows") {
  file_path <- "J:/Summer\ Pathology/Jo\ White/Theta\ Probe\ Files/"
} else # macOS, so
  file_path <-
  "/Volumes/centreforcrophealth/Summer\ Pathology/Jo\ White/Theta\ Probe\ Files/"
}
```

## Using get_soil_moisture() to download new data

This function, `get_soil_moisture()`, downloads data from the FTP server. 
The theta probe logger files are made up of several disjoint files that this
function will concatenate with any prior data files available locally to create
a single master file of the data for cleaning and analysis.

- `userpwd`: provide a valid user name and password to access the FTP server

- `path`: File path to directory for saving a comma separated file (CSV)
#' output. Defaults to current working directory.

- `local_dirs`: File path to directory, which holds previous data logger data
monthly file folders, which contain hourly CSV files from theta probe loggers.

```{r download_data, eval=FALSE}
get_soil_moisture(userpwd = "",
                  path = "~/Documents/Data/Theta Probe/",
                  local_dirs = "~/Documents/Data/Theta Probe/")

```

## Using clean_theta_data() to remove outliers in the data 

Use a Hampel filter to remove outliers before summarising five-minute data into
daily values.

```{r clean_data, eval=FALSE}
soil <- clean_theta_data(csv_in = paste0(
  "~/Documents/Data/Theta Probe/",
  Sys.Date(),
  "_Soil_Moisture.csv"
))
```

## Final formatting and graphing

### Name probes in data file

The soil probes are referred to as "JW_01" and "JW_02" in the logger files. In
this step we will assign the location names in the final file for analysis.

```{r name_probes, eval=FALSE}
soil$Probe[soil$Probe == "JW_01"] <- "Hermitage"
soil$Probe[soil$Probe == "JW_02"] <- "Kingaroy"
```

### Clean up data

The Hermitage probe was reactivated while in storage and transmitted values.
This step removes them from the final data.

```{r remove_extra, eval=FALSE}
soil <-
  soil[which(soil$Probe != "Hermitage" |
               soil$Date < "2017-08-15"), ]
```

## Final steps, graphing and saving data

### Graph data

Create a graph of the data and store it in a shared location. This makes use of
the _theme.usq_ package to apply USQ colours and typography to the graphic.

```{r graph_data, eval=FALSE}
theta_graph <-
  ggplot(soil, aes(x = Date, y = Moisture, colour = Probe)) +
  geom_line() +
  scale_colour_manual(values = usq_palette[c(1, 3)]) +
  ylab("Soil Moisture (%)") +
  facet_grid(Probe ~ .) +
  theme_usq() +
  theme(legend.position = "none")

ggsave(paste0("/Daily_Soil_Moisture_Graph_", Sys.Date(), ".png"),
       theta_graph,
       path = file_path)

```

### Save the cleaned master CSV file

Save data to disk in a shared location.

```{r save_final_data, eval=FALSE}
write_csv(soil,
          paste0(file_path, "Daily_Soil_Moisture_Data_", Sys.Date(), ".csv"))
```

